### 2024-11-01 posts/Algorithmic Nudge/nudge.qmd

---
title: "Algorithmic Nudging: An Application in Curbing Problematic Smartphone Use"
date: 04/16/2024
date-modified: last-modified
date-format: full
categories: [Others]
#image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

# Algorithmic Nudging

An Application in Curbing Problematic Smartphone Use

<iframe src="pdfs/Behavior_Econ_Nudging.pdf" style="width:100%; height:600px;" frameborder="0"></iframe>

You can download it [here](pdfs/Behavior_Econ_Nudging.pdf).