### 2024-12-26 index.qmd
mainfont: Lato monofont: Roboto

---
title: "Shaokang Yang"

---

-   **Ph.D. in Business Information Technology (Expected)**

Virginia Polytechnic Institute and State University

-   **M.Sc. in Data Science and Economics**

University of Milan

-   **Bachelor of Engineering/Bachelor of Economics**

Jimei University