### 2025-01-05 ProjectBlog/paper1/Clustering Anlysis/cluster.qmd
Research blog moving

---
title: "Behavior Segmentation for Smartphone Use "
date: 12/09/2023
date-modified: last-modified
date-format: full
categories: [Paper1]
image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
---

We cluster the smartphone use behaviors based on the three major metrics of smartphone use actions: Screen time, unlock frequency and message.

## Descriptive Statistics

```{python}
#| echo: false
#| warning: false
#| output: true
import time;import pandas as pd;import numpy as np;import datetime
from datetime import datetime;from datetime import timedelta;import sqlite3;import warnings
import matplotlib.pyplot as plt;import seaborn as sns;warnings.filterwarnings('ignore')

df=pd.read_csv("/Users/shaokangyang/Library/CloudStorage/GoogleDrive-sky.ang510@gmail.com/My Drive/Nethealth/Data/CNSA/combine/7/df7_3hr.csv")
df1 = df
df1[['sleep_debt','sleep_duration','screen_time','message','unlock','especiallyclose_m','family_m','friend_m','romantic_m','naps','naps_dur']].describe(percentiles=[.10,.15,.20,.25,.50,.55,.60,.65,.70, .75,.80,.85,.90,.95]).round(2)
```

## Trend Analysis

First, we explore the relationships among variables of sleep debt, screen time (\>15 min) and message with measurement window of three hours:

```{python}
#| echo: false
#| warning: false
#| output: true

# Creating a special bin for 0
df1=df[df['screen_time']>15]
bins = [-np.inf, 0.001] + list(range(15, 185, 5))  # 0.1 to include 0 in the first bin exclusively
df1['screen_time_bin'] = pd.cut(df1['screen_time'], bins=bins, right=False, include_lowest=True)
# Recalculating the aggregated data
grouped = df1.groupby('screen_time_bin').agg({'sleep_debt':'mean', 'message':'mean', 'screen_time':'count'}).reset_index()
# Replotting with the adjusted bins
fig, ax1 = plt.subplots(figsize=(14, 8))
# Line plot for sleep debt trend
line1, = ax1.plot(grouped['screen_time_bin'].astype(str), grouped['sleep_debt'], color='blue', linestyle='-', marker='o', label='Sleep Debt')
# Adding labels for sleep debt
for x, y in zip(grouped['screen_time_bin'].astype(str), grouped['sleep_debt']):
    ax1.text(x, y, f'{y:.2f}', va='bottom', ha='center', color='blue')
# Setting the left y-axis for sleep debt
ax1.set_xlabel('Screen Time (minutes)')
ax1.set_ylabel('Average Sleep Debt')
ax1.set_title('Sleep Debt, Screen Time and Message')
ax1.tick_params(axis='x', rotation=45)
# Line plot for message count on a secondary y-axis
ax2 = ax1.twinx()
line2, = ax2.plot(grouped['screen_time_bin'].astype(str), grouped['message'], color='darkorange', linestyle='--', marker='o', label='Message')
# Adding labels for message count
for i, txt in enumerate(grouped['message']):
    ax2.annotate(f'{txt:.2f}', (grouped['screen_time_bin'].astype(str)[i], grouped['message'][i]), textcoords="offset points", xytext=(0,10), ha='center')
ax2.set_ylabel('Message')
# Bar plot for the number of observations on a secondary y-axis
ax3 = ax1.twinx()
bars = ax3.bar(grouped['screen_time_bin'].astype(str), grouped['screen_time'], width=0.4, color='lightgrey', alpha=0.5, label='Number of Observations')
ax3.spines['right'].set_position(('outward', 60))  # Offset the right spine of ax3
ax3.set_ylabel('Number of Observations')
# Labeling the number of observations above each bar
for bar in bars:
    yval = bar.get_height()
    ax3.text(bar.get_x() + bar.get_width() / 2, yval, f'{int(yval)}', va='bottom', ha='center', color='black')
# Adjusting the plot
ax1.grid(True, which='both', linestyle='--', linewidth=0.5)
ax2.grid(False)
ax3.grid(False)
# Adding legends
fig.legend(loc='upper right')
fig.tight_layout()
plt.show()
```

## Clustering Analysis

### Cluster=11

#### Front view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=11,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#c27ba0',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b',   # Brown
    '#c27ba0'
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 12):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=11)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
plt.show()
```

#### Rear view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=11,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#c27ba0',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b',   # Brown
    '#c27ba0'
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 12):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=11)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
ax.view_init(30, 120)
plt.show()

```

### Descriptive Statistics by Cluster

#### Sleep Debt

```{python}
#| echo: false
#| warning: false
#| output: true

summary = df.groupby('cluster')['sleep_debt'].describe().round(2)
print(summary)
```

The above summary suggests that **cluster 2** has the minimal sleep debt of **31.53** minutes. In comparison, **cluster 10** has the maximal sleep debt of **54.76** minutes.

#### Screen Time

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['screen_time'].describe().round(2)
print(summary)

```

#### Unlock

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['unlock'].describe().round(2)
print(summary)
```

#### Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['message'].describe().round(2)
print(summary)
```

The table above suggests the cluster 2 has the **second highest message frequency**, 79.57. It implies that message might reduce sleep debt.

#### Close Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['especiallyclose_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 2 has the **highest close message frequency**, 49.06. It implies that close message might reduce sleep debt.

#### Family Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['family_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 2 has the **highest family message frequency**, 5.89. It implies that family message might reduce sleep debt.

#### Friend Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['friend_m'].describe().round(2)
print(summary)
```

#### Romantic Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['romantic_m'].describe().round(2)
print(summary)
```

#### Steps

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['steps'].describe().round(1)
print(summary)
```

### Cluster=10

#### Front view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=10,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 11):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=10)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
plt.show()
```

#### Rear view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=10,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 11):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=10)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
ax.view_init(30, 60)
plt.show()
```

### Descriptive Statistics by Cluster

#### Sleep Debt

```{python}
#| echo: false
#| warning: false
#| output: true

summary = df.groupby('cluster')['sleep_debt'].describe().round(2)
print(summary)
```

The above summary suggests that **cluster 6** has the minimal sleep debt of **31.41** minutes. In comparison, **cluster 4** has the maximal sleep debt of **54.50** minutes.

#### Screen Time

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['screen_time'].describe().round(2)
print(summary)

```

#### Unlock

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['unlock'].describe().round(2)
print(summary)
```

#### Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['message'].describe().round(2)
print(summary)
```

The table above suggests the cluster 4 has the **second highest message frequency, 78.15**. It implies that message might reduce sleep debt.

#### Close Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['especiallyclose_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 6 has the **highest close message frequency, 47.81**. It implies that close message might reduce sleep debt.

#### Family Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['family_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 6 has the **highest family message frequency, 5.83**. It implies that family message might reduce sleep debt.

#### Friend Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['friend_m'].describe().round(2)
print(summary)
```

#### Romantic Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['romantic_m'].describe().round(2)
print(summary)
```

#### Steps

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['steps'].describe().round(1)
print(summary)
```

### Cluster=9

#### Front view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=9,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 10):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=9)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
plt.show()
```

#### Rear view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=9,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 10):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=9)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
ax.view_init(30, 60)
plt.show()
```

### Descriptive Statistics by Cluster

#### Sleep Debt

```{python}
#| echo: false
#| warning: false
#| output: true

summary = df.groupby('cluster')['sleep_debt'].describe().round(2)
print(summary)
```

The above summary suggests that **cluster 5** has the minimal sleep debt of **31.47** minutes. In comparison, **cluster 8** has the maximal sleep debt of **54.61** minutes.

#### Screen Time

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['screen_time'].describe().round(2)
print(summary)

```

#### Unlock

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['unlock'].describe().round(2)
print(summary)
```

#### Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['message'].describe().round(2)
print(summary)
```

The table above suggests the cluster 5 has the **second highest message frequency, 77.59**. It implies that message might reduce sleep debt.

#### Close Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['especiallyclose_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 5 has the **highest close message frequency, 47.44**. It implies that close message might reduce sleep debt.

#### Family Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['family_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 5 has the **highest family message frequency, 5.76**. It implies that family message might reduce sleep debt.

#### Friend Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['friend_m'].describe().round(2)
print(summary)
```

#### Romantic Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['romantic_m'].describe().round(2)
print(summary)
```

#### Steps

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['steps'].describe().round(1)
print(summary)
```

### Cluster=8

#### Front view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=8,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 9):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=8)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
plt.show()
```

#### Rear view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=8,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 9):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=8)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
ax.view_init(30, 60)
plt.show()

```

### Descriptive Statistics by Cluster

#### Sleep Debt

```{python}
#| echo: false
#| warning: false
#| output: true

summary = df.groupby('cluster')['sleep_debt'].describe().round(2)
print(summary)
```

The above summary suggests that **cluster 5** has the minimal sleep debt of **31.47** minutes. In comparison, **cluster 8** has the maximal sleep debt of **54.61** minutes.

#### Screen Time

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['screen_time'].describe().round(2)
print(summary)

```

#### Unlock

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['unlock'].describe().round(2)
print(summary)
```

#### Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['message'].describe().round(2)
print(summary)
```

The table above suggests the cluster 5 has the **second highest message frequency, 77.59**. It implies that message might reduce sleep debt.

#### Close Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['especiallyclose_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 5 has the **highest close message frequency, 47.44**. It implies that close message might reduce sleep debt.

#### Family Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['family_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 5 has the **highest family message frequency, 5.76**. It implies that family message might reduce sleep debt.

#### Friend Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['friend_m'].describe().round(2)
print(summary)
```

#### Romantic Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['romantic_m'].describe().round(2)
print(summary)
```

#### Steps

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['steps'].describe().round(1)
print(summary)
```

### Cluster=7

#### Front view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=7,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 8):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=7)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
plt.show()
```

#### Rear view

```{python}
#| echo: false
#| warning: false
#| output: true

from sklearn.cluster import KMeans
from sklearn.preprocessing import StandardScaler
import matplotlib.pyplot as plt
from mpl_toolkits.mplot3d import Axes3D
random_seed = 1
#df1 = df[df['message'] < 240]
df1=df
df1.loc[df1['unlock'] > 41, 'unlock'] = 41;df1.loc[df1['message'] > 96, 'message'] = 96
# Normalizing the data
scaler = StandardScaler()
df_scaled = scaler.fit_transform(df1[['screen_time', 'unlock', 'message']])
# KMeans clustering on normalized data
kmeans = KMeans(n_clusters=7,random_state=random_seed)
clusters = kmeans.fit_predict(df_scaled)
# Adding the cluster data to the original dataframe
df1['cluster'] = clusters+1
# Visualizing the clusters in 3D with normalized data
# Custom colors for each cluster (12 in total)
nature_colors = [
    '#1f77b4',  # Blue
    '#ff7f0e',  # Orange
    '#2ca02c',  # Green
    '#d62728',  # Red
    '#9467bd',  # Purple
    '#ffbb78',  # Light Orange
    '#7a876f',  # Pink
    '#17becf',  # Gray
    '#bcbd22',  # Olive
    '#FFC300',  # Yellow
    '#aec7e8',  # Light Blue
    '#8c564b'   # Brown
]
fig = plt.figure(figsize=(16, 9.6))
ax = fig.add_subplot(111, projection='3d')
# Plotting each cluster with a custom color
for i in range(1, 8):  # Cluster numbers now range from 1 to 12
    cluster_data = df1[df1['cluster'] == i]
    ax.scatter(cluster_data['screen_time'], cluster_data['unlock'], cluster_data['message'], 
               color=nature_colors[i - 1], label=f'Cluster {i}', marker='o')
ax.set_title('Behavior Segmentation for Smartphone Use (Cluster=7)')
ax.set_xlabel('Screen Time (minutes)')
ax.set_ylabel('Unlock Frequency')
ax.set_zlabel('Number of Messages')
# Legend
ax.legend(title="Clusters")
ax.view_init(30, 60)
plt.show()

```

### Descriptive Statistics by Cluster

#### Sleep Debt

```{python}
#| echo: false
#| warning: false
#| output: true

summary = df.groupby('cluster')['sleep_debt'].describe().round(2)
print(summary)
```

The above summary suggests that **cluster 3** has the minimal sleep debt of **35.79** minutes. In comparison, **cluster 6** has the maximal sleep debt of **53.48** minutes.

#### Screen Time

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['screen_time'].describe().round(2)
print(summary)

```

#### Unlock

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['unlock'].describe().round(2)
print(summary)
```

#### Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['message'].describe().round(2)
print(summary)
```

The table above suggests the cluster 3 has the **a message frequency of 7.56**. It implies that message level is not necessarily associated with sleep debt level.

#### Close Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['especiallyclose_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 3 has the **a close message frequency of 2.90**. It implies that close message level is not necessarily associated with sleep debt level.

#### Family Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['family_m'].describe().round(2)
print(summary)
```

The table above suggests the cluster 3 has the **a family message frequency of 0.74**. It implies that family message might reduce sleep debt.

#### Friend Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['friend_m'].describe().round(2)
print(summary)
```

#### Romantic Message

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['romantic_m'].describe().round(2)
print(summary)
```

#### Steps

```{python}
#| echo: false
#| warning: false
#| output: true
summary = df.groupby('cluster')['steps'].describe().round(1)
print(summary)
```