### 2025-01-05 posts.qmd

---
title: "Project Blog"

website:
  favicon: Virginia_Tech_seal.svg.png
  image: Horizontal_VT_Full_Color_RGB.png
  sidebar:
    tools:
      - icon: github
        href: https://github.com/Shaokang-Yang
      - icon: linkedin
        href: https://www.linkedin.com/in/shaokang-y-52b633162/
    section: true
    contents:
      - href: index.qmd
        text: Home
      - research.qmd
      - teaching.qmd
      - posts.qmd
---