### 2025-01-06 ProjectBlog/paper1/Hypotheses/hypo.qmd
delete posts

---
title: "Review for Production and Operations Management(2024)"
date: 09/16/2024
date-modified: last-modified
date-format: full
categories: [Others]
#image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

# Review for Production and Operations Management(2024)

The latest review report for the manuscript **Breaking Barriers: Improving Patient Adherence to Appointments and Provider Productivity through Telehealth(POM-Jul-24-OA-1071)** is as follows:

<iframe src="pdfs/Reviewer_Report_POM_2024.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>