### 2025-01-06 posts/Paper 1/paper1.qmd

---
title: "Smartphone Use, Social Support, and Sleep Health"
date: 10/09/2023
date-modified: last-modified
date-format: full
categories: [Paper1]
image: "Go2Bed-02.png"
format: revealjs
---

## Falling Asleep Hours Distribution

![](images/Go2Bed-02.png){fig-align="center"}

Normal Sleep: fall asleep from 22 pm to 6 am. (**95.99** % of Total.)

## Falling Wake Hours Distribution

![](images/wake.png){fig-align="center"}

## Interaction of Time terms

::: incremental
-   **Insession** x **Day of Week** x **Go2Bed Hour**:\
    1 x 5 x 2: [2 am on Friday of a Semester]{.underline}
-   Abnormal Screen Time Threshold: 12 -\> 3.5 hours
-   Screen time as control: We focus on **Technology** impact under the influence of **Human** factors
:::

## Conclusion

::: incremental
-   Social interaction via smartphone can **benefit** sleep quality
-   Interaction with **close** contacts prolong sleep
-   Interaction with certain **relationships** of contacts (e.g. family) prolong sleep
:::

## Moderating Impact: The Way of Interaction

::: incremental
-   Message: **Positive** moderation
-   Call: Depends on the **relationship** and **closeness**
:::

## Result 1: Social Closeness via Message

![](images/MsgClose.png){fig-align="center"}

## Result 2: Social Closeness via Call

![](images/Close_Cal.png){fig-align="center"}

## Result 3: Relationship via Message

![](images/Rel_Msg.png){fig-align="center"}

## Result 4: Relationship via Call

![](images/Rel_Cal.png)

## Result 5: Pre-Sleep Communication benefit Sleep Quality

![](images/ComCM.png){fig-align="center"}