### 2025-01-06 teaching.qmd

---
title: "Teaching"
format:
  html:
    css: styles2.css
class: custom-font-pos
---