### 2025-01-06 posts/SP2 Review Response/ReviewResponse.qmd
delete project text

---
title: "SP2 Reveiw Response: Social Network Activation: \\The Application in Smartphone Use"
date: 12/23/2023
date-modified: last-modified
date-format: full
categories: [Paper2]
image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
bibliography: references.bib
---

We thank Dr.Adjerid, Dr.Liu, and Dr.Lowry for your insightful and valuable feedback and for the opportunity to submit a revision. In addition to highlighting the areas that need to be improved, all of you gave detailed suggestions and intensive oral/written instructions I might take to address the concerns. I sincerely appreciate your constructive comments and advice. The manuscript has been substantially revised in response to the review team's comments. Below I provide an overview of the main changes.

### **\[Development of Motivation\]**

In revising the manuscript, we conduct a comprehensive literature review about social engagement, social network activation, and smartphone overuse, through which we can better identify the practical problems and gaps in the previous literature and clarify the contributions of this study. Specifically, we summarize three major practical and research motivations in the area.

1.  Practical problem. The literature on social engagement and smartphone overuse suggests that lack of social engagement can lead to negative health outcomes. A primary consequence of social isolation that is increasingly popular in life but can be easily neglected is smartphone overuse. It is important to examine how to strategically activate social ties so that smartphone users can improve their social engagement level and reduce the excessive time spent on smartphones.

2.  Research gaps. The research gaps of the study are reflected in the following aspects:

-   Previous research on social engagement has largely focused on the volume of social interaction, such as the number of contacts and the frequency of interactions, little attention has been given to the impact of social network structure or topology on an individual's behavior and well-being.

-   An equally important, but understudied, aspect of social engagement is dynamic evolutions in engagement with different parts of the social network-- the phenomenon termed social network activation. Most social networks in the current literature possess a static and cross-sectional nature, potentially causing them to overlook the intrinsic dynamics in social interactions. We leverage a longitudinal study to meticulously craft a dynamic social network framework. This innovative approach allows us to meticulously trace the progression of individuals' social networks over time, capturing the essence of their evolving connections and interactions.

In addition to the above modification, we introduce and emphasize the technology components in the process of social network activation to make the manuscript fit IS research journals.

### **\[Theorizing and hypothesis development\]**

As suggested by the review team, we conduct a theoretical development for the network features we extracted (degree centrality and K-core), which allows us to develop three hypotheses. Drawing on the three streams of literature on social engagement and smartphone overuse, we first hypothesize that **activating social networks reduces smartphone screen time; the reduction in screen time increases with the number of contacts activated (H1)**. Furthermore, we identify key nodes with high social influence. We infer that contacts with a high degree centrality have a high capacity to motivate social engagement, therefore, our second hypothesis is: **Activating social networks with a higher score of average degree centrality in a social network reduces the time spent on a smartphone (H2)**. Nevertheless, we speculate that degree centrality cannot represent a contact's comprehensive influence in a social network as it may neglect a contact's indirect influence, which is the capacity to form social groups and mobilize other contacts. Thus, we utilize another essential social network feature, "K-core," to measure a contact's social influence by considering the connectivity of the contact as well as the cohesiveness of the group in which the contact exists. Therefore, we propose our third hypothesis: **Activating social networks with a higher average core number score can reduce the time spent on a smartphone (H3)**.

### \[Analysis\]

Our analysis has been improved in several ways. In addition, multiple robustness checks are conducted to validate our estimation results.

1.  We adjust the threshold of activation. In the first version of this manuscript, the activation threshold is communication event \> 0. However, there could be a situation in which the contacts did not respond to the information, which lead to a failed activation. We adjust the activation threshold to ensure a two-way communication (an effective activation) exists between two smartphone users. This eliminates the possibility of one-way communication: one only sends or receives communication information. Overall, the estimation results are mostly consistent with earlier results after the adjustment.

2.  We identified smartphone users who are addicted to smartphones. To the best of my knowledge, there is no quantitative definition of smartphone addiction. We define smartphone addiction based on multiple criteria. Finally, we conduct additional robustness checks to rule out alternative explanations and to check reverse causality. All estimation results show that the increase in core number is associated with a decrease in the likelihood of smartphone addiction.

3.  We conducted a lookup operation to search for the missing contacts’ social network information from the closest neighbors’ social network surveys. We also conducted an alternative lookup approach by searching for the missing contacts backward first and then forward. In both cases, we borrow the missed contacts’ network features from the closest wave that the contact exists. By doing this, we could have a relatively complete social network.  The estimation with an adjusted approach is mostly consistent with earlier estimations.

### **\[Writting\]**

In an effort to increase the readability of the paper, we have overhauled the structure and logic of the paper, tightened the analysis, and checked all the grammatical errors and citations.

::: panel-tabset
## Dr.Idris Adjerid {#sec-dr.idris-adjerid}

Thank you for your valuable comments and feedback. Please refer to the manuscript for the writing update.

+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| No. | Comments                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | Responses                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |
+=====+====================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================+==============================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================+
| 1   | Does the analysis differ much with and without this approach, could be better as robustness. Could address the issue of up to date social network.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 | Thank you for your constructive comments. Yes, the analysis with this approach and an alternative approach show similar estimation results. The objective of the approach is to search for the missing contacts' social network information from the closest neighbors' social network surveys. This approach starts from a forward lookup followed by a backward lookup for missed contacts in the current wave of social networks. We also conducted an alternative lookup approach by searching for the missing contacts backward first and then forward, which could work as robustness checks.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | ![](images/search.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             | ![](images/search%20-%20bf-01.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | In both cases, we borrow the missed contacts' network features from the closest wave that the contact exists. By doing this, we could have a relatively complete social network.  The estimation with an adjusted approach is **mostly consistent** with earlier estimations:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/clipboard-2450557691.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The results suggest that an increase in core number could lead to a decrease in screen time, which is consistent with the original estimation.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 2   | *"dynamic social network"*                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         | Thank you for pointing that out. It sill should be noted that there are two levels of social network features in our study:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | Actually this definition is not clear. Social networks can be dynamically measured (i.e. they are repeated snapshots). Would these be dynamic? In reality, you are taking a cross-sectional snapshot of a social network and dynamically measuring activation. You actually do not have dynamic social network data.                                                                                                                                                                                                                                                                                                               | 1.  **Individual level**. Contacts' original network features come from the social network survey, therefore, they are updated over the waves of surveys.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | 2.  **Cluster/group level**. Everyday, participants selectively activate a group of individual contacts. The network features of the activated cluster change day by day.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | It is true that individual's social network features evolve survey by survey rather than day by day. Nevertheless, our **independent variables** are the average social network features of the entire **cluster** of contacts that are activated on a certain day. Depending on which contacts are activated, the characteristics of the activated cluster change on daily level. For example, if a participant activate the following contacts on day t:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | Participant | Contacts activated | Date | K-Core (Individual ) | K-Core (Group) |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | |-------------|--------------------|------|----------------------|----------------|                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | S           | A                  | t    | 5                    | 6              |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | S           | B                  | t    | 6                    | 6              |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | S           | C                  | t    | 7                    | 6              |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The K-core for S on day t is (5+6+7)/3=6. On day t+1, the participant activate the following contacts:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | Participant | Contacts activated | Date | K-Core (Individual ) | K-Core (Group) |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | |-------------|--------------------|------|----------------------|----------------|                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | S           | C                  | t+1  | 7                    | 4              |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | S           | E                  | t+1  | 2                    | 4              |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | | S           | H                  | t+1  | 3                    | 4              |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The K-core for S on day t+1 is (7+2+3)/3=4. As we can see, the features of contacts A, B, C, E, H does not change in the current wave of survey. However, the feature of the activated cluster comprised of those contacts sill vary day by day.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | In general, I can feel that the terminology "dynamic social network" could still be controversial or inaccurate, therefore, I have replaced it with "social network" in the manuscript for now.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 3   | *" Moreover, it helps us capture the evolution of one’s contact list and their position in the network over time, which is the common nature of one’s social interaction. "*                                                                                                                                                                                                                                                                                                                                                                                                                                                       |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | In some sense, the social network measurement is dynamic both in the evolving structure and the activation. These both need to be highlighted and cleanly separated.                                                                                                                                                                                                                                                                                                                                                                                                                                                               |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 4   | *"smartphone addiction"*                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | Thanks for your correction, since there is no universal or quantitative definition for smartphone addiction, I have used the term "smartphone overuse" to substitute "smartphone addiction" in the manuscript.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | Unless we define this (super clearly it could be confusing, we may want to use another, less loaded term (e.g. “phone overuse”                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 5   | *"Our findings reveal that activating contacts who are embraced by the other contacts in one’s social network helps reduce smartphone use. The findings advance the understanding of social engagement from volume to the structure of social networks."*                                                                                                                                                                                                                                                                                                                                                                          | Thanks for your comments.  I modified the statement as follows by adding an explanation of the results:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | How to interpret these results is still unclear, need to have a coherent explanation of results and why these would or would not be expected.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      | Our findings reveal that the topology or structure of one's social network impacts individuals' behavior. Specifically, activating contacts embraced by the community of one’s social network **could motivate broader social engagement**, which helps curb smartphone overuse. Our study advances the understanding of social engagement from volume to the structure of social communities.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 6   | *"......we contextualize the research prob- lems in the context of smartphone addiction, which is an emerging social and technical prob- lem and is broadly thought to be associated with the lack of social engagement."*                                                                                                                                                                                                                                                                                                                                                                                                         | Thank you for your comments. The statement has been corrected as:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                            |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | Here there is this argument again. I’m not sure what we are saying. At points it sounds like we are saying that smartphone addiction is related to social isolation. Alternatively, social isolation could lead to smartphone overuse, both seem credible but the latter is more sensical for our story so we should be clear about this throughout the paper.                                                                                                                                                                                                                                                                     | "......we narrow down the research questions to smartphone overuse, which is an emerging social and technical problem that is broadly thought to be a consequence of social isolation[@zeng2022]".                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 7   | *" ......we measure a social network member’s core number, which measures how well a contact is connected with other contacts and how active the contact’s community is."*                                                                                                                                                                                                                                                                                                                                                                                                                                                         | Thank you for your thorough advice. I think the following theoretical construct could be related to our study:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | This is crucial and should be more clearly described. What about centrality? There is a key theoretical construct here which is measured in different ways. So we should identify the key construct and how the two measures we have differ (and by extension setting up why they may different effects).                                                                                                                                                                                                                                                                                                                          | 1.  Social influence. The construct reflects the capacity of certain nodes in a social network to affect others' behavior, opinion, and decisions-making process [@izuma2017]. (Social influence. could be a construct for degree centrality and K-Core)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | 2.  Group cohesion. The construct describes to which degree group members are attracted and connected in their group. Group cohesion can be based on shared goals or interests, emotion support, and group pride. Evidence suggests that group cohesion is positively correlated with team performance [@paulus2011]. (The constructs fits better for K-Core)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | I take "social influence" as the theoretical construct for now. The statement is modified as follow:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | *In addition to the fundamental network metrics such as size and number of edges, we utilize degree centrality and K-core to identify the vital contacts with high social influence in a network that could have the potential to motivate broader social engagement. The former feature captures the degree to which a contact itself is connected to other contacts. Meanwhile, the latter further considers the connectivity degree of the sub-network to which a contact belongs.*                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | I wonder if we should identify one theoretical construct for both degree centrality and K-core. Or, should we use two theoretical constructs for degree centrality and K-core separately.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 8   | *"However, the number of connections of an activated network does not contribute to curbing smartphone use. Moreover, our results suggest that activating contacts in one’s social network can curb smart- phone use."*                                                                                                                                                                                                                                                                                                                                                                                                            | This interpretation is not so clear. Here is the more detailed interpretation of the current results:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | These seem like opposite results, are they? Is the first result suggesting that network size does not have an effect? More generally, I wonder if we should kick things off with the more basic measures and then lead into the new more comprehensive measures. The story is that social networks are rich constructs with some simpler measures (number of nodes, size, etc.) and some more advanced measures that capture richer dynamics. Part of the contribution is leaning into the richer context and using real time data reveal more insight compared to the simple stuff. Need to tell this story more comprehensively. | Our results show that activating contacts with high K-core can significantly reduce smartphone overuse time, with a more substantial reduction effect on non-communicative use than communication-driven use. Specifically, one standard deviation increase in core number leads to a 5.7-minute decrease in non-communication screen and a 5.34% decrease in overuse rate. However, activating contacts with a high degree centrality does not help curb smartphone overuse. The distinction between the effects of the two features suggests that contacts with a high K-core, which are those embedded in a cohesive group, have a stronger capability to motivate social engagement and project social influence. Furthermore, we find that the size of an activated social network reduces non-communicative screen time but increases communicative screen time. The two types of effects offset each other and result in no significant impact on overall screen time. The two types of effects offset each other and result in no significant impact on overall screen time. Moreover, we find that the impact of social network features does not vary with mobile communication level. Our results reveal that once a cluster of social networks is activated, the communication level does not influence the impact of the activated cluster, even though the activation itself is based on communication events. |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 9   | *"There are generally two types of social isolation: social disconnectedness and perceived isolation. Social disconnectedness refers to the status that someone lacks contacts or social con- nections. "*                                                                                                                                                                                                                                                                                                                                                                                                                         | It is interesting to explore the distinction between social disconnectedness and perceived isolation. In our data, there are three types of perceived loneliness: romantic loneliness, family loneliness and social loneliness. We can use the number of activated contacts per day as a proxy of social disconnectedness. Here is some descriptive results:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     | These are interesting constructs which waiver in their centrality to the story. Do they correlate with our social network measures, are there other measures that capture these distinctions?                                                                                                                                                                                                                                                                                                                                                                                                                                      | The correlation between family loneliness and social disconnectedness:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/paste-8.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/paste-2.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The figure suggests that high family loneliness is correlated with more screen time regardless of the number of contacts activated.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The correlation between romantic loneliness and social disconnectedness:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/paste-9.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/paste-4.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The figure suggests that low romantic loneliness is correlated with high screen time regardless of the number of contacts activated.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The correlation between social loneliness and social disconnectedness:                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/paste-10.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | ![](images/paste-3.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | The figure suggests that high social loneliness is correlated with high screen time when participants' activated contacts are at bottom 20% or top 20%.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: Review Response for Dr.Idris Adjerid

## Dr.Jiayi Liu {#sec-dr.jiayi-liu}

Thank you for your valuable comments and feedback.

+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| No. | Comments                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         | Responses                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
+=====+==================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================+===========================================================================================================================================================================================================================================================================================================================================================================================================================================================================================================+
| 1   | It appears to be that those with higher core-number are mostly close friends. Then one could say that the story is actually much simpler: merely engaging with close friends is what matters. Can you argue for using network features, like core-number, against more intuitive metrics such as contact closeness in this context?                                                                                                                                                                                                                                                                              | Thank you for your constructive comments. We measured the social distance of smartphone users with their contacts, as well as their level of trust in those contacts, which are the common measurement of interpersonal relationships.                                                                                                                                                                                                                                                                    |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | The estimation of the above two constructs suggest that they do not have significant impact on participants smartphone use behavior. Please refer to Table 5:                                                                                                                                                                                                                                                                                                                                             |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | ![](images/close_trust.png){width="35cm"}                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 2   | A fundamental assumption is that social interactions mainly occur through calls and messages. Although the author has presented data showing a link between self-reported interaction levels and observed call/message frequencies, this doesn't establish that calls/messages are the primary channels of interaction. People could have spent more time on other social platforms and offline activities (with the time spent also aligning with the self-reported interaction levels). How do you plan to address these potential unobserved interactions? Or can you justify they don't affect your results? | This is an important issue that we need to address. Thanks for pointing this out. Indeed, there could be unobservable social engagements that are not based on phone call and message: 1) social interactions on other mobile applications; 2) physical interactions that are not based on phone call or messages.                                                                                                                                                                                        |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | First, for the social interactions happening on other platforms, it is indeed unobservable in our data set. However, based on the market share of the most popular mobile messaging channels, our data covered the leading four channels (Whatsapp, Phone calls, Facebook messager and Texting (SMS):![](images/channel.png)                                                                                                                                                                              |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | Therefore, we suppose that our data has covered the majority of mobile communication activities.                                                                                                                                                                                                                                                                                                                                                                                                          |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | Second, in terms of the unobserved offline social interactions, there could be two categories:                                                                                                                                                                                                                                                                                                                                                                                                            |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 1.  Smartphone users' offline social interactions were initiated by online communication and then it was swithed to offline social interaction. In this situation , the activation of social network is activated by smartphone communication. This scenario could support our 1 & and 2 hypothesis: activating embedded contacts and interacting with them could facilitate offline social engagement.                                                                                                   |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 2.  The primary concern is the **offline social interaction is independent of smartphone communication**. In another word, the activation of social network is not based on smartphone communication. Nevertheless, 67.91% (90134/132709) of our observations are in session, which means students are on campus. For those observations, unobserved social interactions within family are ruled out. We speculate that students started with online social engagement except those with their roommates. |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 3   | The implications from your findings are also not clear. The current results show reduced smartphone use when engaging high "core-number" contacts. Does this suggest that one should have more interactions with such contacts to minimize their screen time? But the effect seems to be muted for the heaviest smartphone users, meaning this strategy wouldn't be effective for those who might be really addicted.                                                                                                                                                                                            | We identified smartphone users who are addicted to smartphone. To the best of knowledge, there is no quantitative definition of smartphone addiction. We define smartphone addiction based on multiple criterias:                                                                                                                                                                                                                                                                                         |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 1.  Observations with daily screen time that are greater than 5 hours (300 minutes).                                                                                                                                                                                                                                                                                                                                                                                                                      |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 2.  Observations with daily screen time that are ranked top 10%, 20%, 30% among population.                                                                                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 3.  Top 10 % screen time observations for each individuals' observations.                                                                                                                                                                                                                                                                                                                                                                                                                                 |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | We then estimated the probability of addiction with a logit model:                                                                                                                                                                                                                                                                                                                                                                                                                                        |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | $\begin{align} \ln \left(\frac{\operatorname{Pr}\left(\text { Addiction }_{i t}=1 \mid \boldsymbol{X}_{i t}\right)}{1-\operatorname{Pr}\left(\text { Addiction }_{i t}=1 \mid \boldsymbol{X}_{i t}\right)}\right) &= \nonumber \\ &\quad \beta_{0} + \beta_{1} \cdot \text{Social Network Features}_{it} \nonumber \\ &\quad+ \alpha \cdot X_{it} + \theta_{i} +\lambda_{t} + \varepsilon_{it}\end{align}$                                                                                                |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | All estimation results show that the increase in core number is associated with a decrease in the likelihood of smartphone addiction. For example:                                                                                                                                                                                                                                                                                                                                                        |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | ![](images/clipboard-2847711747.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 4   | Does a single interaction by call/text suffice to trigger this effect?                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | **In the first version of this paper, the activation threshold is communication event \> 0. However, there could be a situation that the contacts did not respond to the information, which lead to a failed activation. I adjust it as follows: a contact is activated if any of the two conditions is met:**                                                                                                                                                                                            |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     | Can you provide more thoughts on the results?                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | 1\) the communication event type is a Call;                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 2\) the communication event type is a message (SMS, MMS, or WhatsApp) and there are both '**outgoing**" and'**incoming**" messages (SMS, MMS, or WhatsApp).                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | The above two conditions could ensure a **two-way communication (an effective activation)** exists between two smartphone users. This eliminates the possibility of one-way communication: one only sends or receives communication information. Overall, the estimation results are **mostly consistent with earlier results** after the adjustment.                                                                                                                                                     |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | ![](images/clipboard-2882367026.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 5   | It could be also interesting to explore how this effect vary by individual's perceived loneliness, their social interaction level, or maybe the timing of activation (e.g., morning vs night), etc.                                                                                                                                                                                                                                                                                                                                                                                                              | The third hypothesis is adjusted as follow:                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | An increase in communication with contacts with a high degree centrality improves the effectiveness of degree centrality in reducing smartphone usage.                                                                                                                                                                                                                                                                                                                                                    |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 6   | Smartphone addiction is measured by duration of smartphone use. But certain usages like seeking information or connecting can be productive. Why not consider defining addiction as screentime over a certain threshold?                                                                                                                                                                                                                                                                                                                                                                                         | Please refer to response 3.                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 7   | "Activation" of a contact in a social media is determined by a single message or call. Shouldn't the frequency of interactions be considered? Would a weighted average make more sense here?                                                                                                                                                                                                                                                                                                                                                                                                                     | In robustness check, we adjust the activtion threshold as two-way communication. Please refer to response 4.                                                                                                                                                                                                                                                                                                                                                                                              |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 8   | In Figure 5, you handle missing contact information by referencing from the past or future. How often does such missing information happen? Wouldn't it be more appropriate to rely on the most recent survey only in order to capture the dynamic relationship?                                                                                                                                                                                                                                                                                                                                                 | We conducted a lookup operation to search for the missing contacts' social network information from the closest neighbors' social network surveys.                                                                                                                                                                                                                                                                                                                                                        |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | ![](images/search.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | We also conducted an alternative lookup approach by searching for the missing contacts backward first and then forward.                                                                                                                                                                                                                                                                                                                                                                                   |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | ![](images/search%20-%20bf-01.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                        |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | Both results are consistent with our original analysis:                                                                                                                                                                                                                                                                                                                                                                                                                                                   |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | ![](images/clipboard-2450557691.png)                                                                                                                                                                                                                                                                                                                                                                                                                                                                      |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 9   | Following the above point, how did you handle new relationships that emerged between biannual surveys?                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | Please refer to response 8.                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 10  | Figure 7 exemplifies social network activation. I'm curious about the repeated communications with a few out-of-network contacts: are these inbound or outbound? If these are mostly inbound with no response or occurred at regular times, it might be more reasonable to classify them as potential ads/promotions and exclude them from analysis.                                                                                                                                                                                                                                                             | 1.  Non-human contacts (promotions, advertising, fishing) are ruled out.                                                                                                                                                                                                                                                                                                                                                                                                                                  |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | 2.  We conducted two-way activation threshold to ensure that activation of social network is successful.                                                                                                                                                                                                                                                                                                                                                                                                  |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 11  | Personality was included as a control. Shouldn't it remain consistent for an individual? If so, you should drop it from the fixed-effect model.                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | Personalities are measured based on social surveys so they could change over surveys. Due to the missing information for certain participants, we excluded the information for now.                                                                                                                                                                                                                                                                                                                       |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 12  | It's not necessary to write three functions for total, communication, and non-communication screen time; the model is the same.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | Thank you, it is corrected in the manuscript.                                                                                                                                                                                                                                                                                                                                                                                                                                                             |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 13  | Based on the \# observations, the unit of analysis seems to be person-week, but in your model and results, "t" indicates day. Please clarify.                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | Thank you, the structure of the data is individual \* day. However, time fixed effect is at the **week level** (**`studyweeks`**), even though our panel data is structured at potentially a more granular (daily) level with **`date`**                                                                                                                                                                                                                                                                  |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 14  | Your hypothesis 3 is about perceived loneliness but the results are missing.  What is the relationship between the perceived loneliness and one's actual communication pattern?                                                                                                                                                                                                                                                                                                                                                                                                                                  | The third hypothesis is adjusted as follow:                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |
|     |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  | An increase in communication with contacts with a high degree centrality improves the effectiveness of degree centrality in reducing smartphone usage.                                                                                                                                                                                                                                                                                                                                                    |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 15  | Please make the tables/figures self-explanatory. For example, in Table 6 it's unclear which column corresponds to which subsample. In Table 4, please specify the units for each variable.                                                                                                                                                                                                                                                                                                                                                                                                                       | Thank you, please refer to the manuscript for the recent correction in formats.                                                                                                                                                                                                                                                                                                                                                                                                                           |
+-----+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: Review Response for Dr.Jiayi Liu
:::