### 2025-01-06 ProjectBlog/paper1/screen time/screentime.qmd

---
title: "RQ2: Effects of Communication-Driven Use"
date: 05/21/2024
date-modified: last-modified
date-format: full
categories: [Paper1]
image: "images/msg_level-mean_winsor_screen.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    toc: true
    toc-location: right
    page-layout: full
bibliography: references.bib
---

# Hypothesis Development

For the hypotheses development, I referred to the  paper  “Thinking about U: Theorizing and testing U- and inverted U-shaped relationships in strategy research” [@haans2016]to motivate and develop our paper. he measure- ment of smartphone usage can be measured by:

1.  Duration

2.  Frequency [@marciano2022]

## The Effects of Overall Smartphone Usage

### Positive Effects (Increasing pathway)

1.  Crowding-out effect on sleep time

2.  Increases exposure time to blue light

### Negative Effects (Decreasing pathway)

Using smartphones for leisure activities increases psycho- logical detachment from work/study [@hülsheger2014]. In addition, “The dose makes the poison.” The impact of overall smartphone use on sleep debt may depend on the amount of use, and moderate smartphone use could benefit sleep:

![](images/U-Shape.png){fig-align="center"}

We hypothesize that:

**Hypothesis 1:** There exists a U-shaped relationship between sleep debt and the amount of smartphone use before sleep.

## Moderating Effects of Social Interaction Usage

We measure social interaction usage by two indicators: message count and network size (contact count). We create a moderator Message Level: The message level is created based on the mean of message frequency(1 if \> mean, 0 if $<$ mean). Same for Network Size Level.

[![Topologies of Socialization](images/clipboard-1010489237.png){fig-alt="Source: flickr" fig-align="center"}](https://www.flickr.com/photos/joelogs/89627794/)

[![Technology and Relationships](images/Relationship_Tech.png){fig-align="center"}](https://andrewsobel.com/article/technology-and-relationships-where-should-you-spend-your-time/)

Based on the topologies about interaction and intimacy, we propose that:

![](images/clipboard-3692274076.png){fig-align="center"}

We estimate the moderating effects of message level and network size with the following specifications:

![](images/clipboard-3930394300.png){fig-align="center"}

The estimation results are as follow:

![](images/clipboard-681847467.png){fig-align="center"}

### Social Interaction as Moderator

![](images/message_moderator.png){fig-align="center"}

### Social Intimacy as Moderator

![](images/Intimacy_moderator.png){fig-align="center"}

### Social Intimacy as Moderator

![](images/Socialization_Moderator.png){fig-align="center"}

## Robustness Checks

![](images/clipboard-2673833856.png)

### ![](images/clipboard-874171734.png)

### ![](images/clipboard-2840737469.png)

### Contact Level as Moderator

![](images/contact_moderator.png){fig-align="center"}

### InnerContact Level as Moderator

![](images/InnerContact-Moderator.png){fig-align="center"}

### Socialization as Moderator

![](images/AlterSocialization_Moderator.png){fig-align="center"}

### Family as Moderator

![](images/contact_family50mp.png)

# Model and Results

## Descriptive Statistics

```{python}
#| echo: false
#| warning: false
#| output: true
import time;import pandas as pd;import numpy as np;import datetime
from datetime import datetime;from datetime import timedelta;import sqlite3;import warnings
import matplotlib.pyplot as plt;import seaborn as sns;warnings.filterwarnings('ignore')

df=pd.read_csv("/Users/shaokangyang/Library/CloudStorage/GoogleDrive-sky.ang510@gmail.com/My Drive/Nethealth/Data/CNSA/combine/7/df7_3hr.csv")

df[['sms','mms','whatsapp','message','length','bytes','screen_time','screen_com','screen_nocom','sc_com_pct','screen_com_level2','com_level']].describe(percentiles=[.10,.15,.20,.25,.50,.55,.60,.65,.70, .75,.80,.85,.90,.95]).round(2)
```

1.  **sms:** SMS message

2.  **mms:** MMS message

3.  **whatsapp:** Whatsapp message

4.  **message:** sms+mms+whatsapp

5.  **length:** Total Length of message in characters within 3 hours

6.  **bytes (KB):** Total Size of image (MMS) in bytes within 3 hours

7.  **screen_time:** overall screen time (minutes) within 3 hours

8.  **screen_com:** communication screen time (minutes) within 3 hours

9.  **screen_nocom:** non-communication screen time (minutes) within 3 hours

10. **sc_com_pct:** percentage of communication screen time (screen_com/screen_time, 0-100)

11. **screen_com_level2:** dummy variable for communication screen time (0: sc_com_pct\<50 %; 1: sc_com_pct ≥ 50 %)

12. **com_level:** dummy variable for message frequency (0: message \<average message ; 1: message ≥ average message)

Since the communication screen time depends on SMS messaging time and MMS messaging time. We calculate the approximate communication time as follow:

![](images/clipboard-3772066647.png){fig-align="center"}

SMS messaging time depends on

-   the text characters (length),

-   SMS texting speed (the number of characters one can text per minute),

-   SMS (the number of SMSs within 3 hours).

MMS messaging time depends on

-   the size of MMS (Bytes),

-   MMS messaging speed (the Bytes one can communicate per minute),

-   MMS (the number of MMSs within 3 hours).

## Communication Vs No-Communication (with Assumptions)

### Impact of Overall Screen Time

![](images/clipboard-1725006988.png){fig-align="center"}

The predicted sleep debt is based on the third column (highlighted), the turning pint for the overall screen time is **90.348** minutes.

![](images/overall_10-01.png){fig-align="center"}

### Comm. vs Non-Comm.

![](images/clipboard-1663320982.png){fig-align="center"}

### Predicted Sleep Debt by NoComm. Screen Time (Interval = 10 min)

![](images/clipboard-3848216028.png){fig-align="center"}

### Predicted Sleep Debt by Comm. Screen Time(Interval = 10 min)

![](images/clipboard-2518025618.png){fig-align="center"}

![](images/ComNocom-02.png){fig-align="center"}

We observe that:

1.  The turning pint of communication screen time comes **later** than that of non-communication screen time, which suggests that communication screen time becomes problematic later than that of non-communication screen time.

2.  The predicted sleep debt of communication screen time is **lower** than that of non-communication screen time

## Moderating Effect of Communication (without Assumptions)

In the the second column, we add an interaction term Message(High/Low) for screen time variables. The predicted sleep debt under low and high message levels are shown as follow:

![](images/msg_level-mean_winsor_screen.png){fig-align="center"}

We observe that:

1.  The turning pint of screen time with **high** message level **later** than that with **low** message level, which suggests that screen time with **high** message level becomes problematic later than that with **low** message level.

2.  The predicted sleep debt of screen time with **high** message level is **lower** than that with **low** message level.

The message level is created based on the mean of message frequency(1 if \> mean, 0 if \< mean). We can observe that message level moderates the impact of screen time by

1) **moving the turning point right**, and

2) **flattening the curve,** which are the two ways of moderation effects in U-shape effects [@haans2016].

## Appendix

### Predicted Sleep Debt by Screen Time Intervals

![](images/clipboard-2036136209.png)

### Predicted Sleep Debt with Moderating Effects by Screen Time Intervals

![](images/clipboard-1013317902.png)

![](images/clipboard-2164487531.png)