### 2025-01-06 ProjectBlog/others/Natality/natality.qmd
natality

---
title: "Dobbs' Decision And Abortion"
date: 01/05/2025
date-modified: last-modified
draft: false
date-format: full
categories: [Others]
website:
  comments: 
    utterances:
      repo: quarto-dev/quarto-web
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
bibliography: references.bib
---

## Background

Restrictions on abortion have rendered significant impact on abortion decision and the number of births. Texas launched two abortion policies:

Senate Bill 8 of Texas is a bill that forbidden abortions for pregnancy over six weeks.