### 2025-01-07 ProjectBlog/others/Natality/natality.qmd
sdid

---
title: "Dobbs' Decision And Abortion"
date: 01/05/2025
date-modified: last-modified
draft: false
date-format: full
categories: [Others]
website:
  comments: 
    utterances:
      repo: quarto-dev/quarto-web
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
    default-image-extension: svg
bibliography: references.bib
---

## Background

Restrictions on abortion have rendered significant impact on abortion decision and the number of births. For example, Texas launched two abortion policies:

-   Senate Bill 8 of Texas: The bill launched on **September 1st, 2021** forbidden abortions for pregnancy over six weeks.

-   Dobbs' decision: The decision passed on **June 24, 2022** allows states to make their own abortion policies.

Any impact on the number of abortions will be captured by the number of births around ten months later.

## Methodology

### Data

County by month level natality observations.

### Model

As the parall trend assumption is hard to satisfy in this study, I made a synthetic Texas with states (e.g., New York and California) with abortion free policies and take it as control group. I use a synthetic difference-in-difference estimator [@arkhangelsky2021] to examine the effects of the two policies on abortions. The matrix for the estimator is as follows:

$$
Y = \begin{bmatrix}
    \pmb{Y}_{pre, co} & \pmb{Y}_{pre, tr} \\
    \pmb{Y}_{post, co} & \pmb{Y}_{post, tr} \\
\end{bmatrix}
$$

Recall the Diff-in-Diff estimator:

$$
Y_{it} = \beta_0 + \beta_1 Post_t + \beta_2 Treated_i + \beta_3 Treated_i  Post_t + e_{it}
$$

we fit unit ($\alpha_i$) and time ($\beta_t$) averages, alongside the treatment indicator.

$$
\hat{\tau}^{did} = \underset{\mu, \alpha, \beta, \tau}{argmin} \bigg\{ \sum_{i=1}^N \sum_{t=1}^T \big(Y_{it} - (\mu + \alpha_i + \beta_t + \tau D_{it}\big)^2 \bigg\}
$$

and the Synthetic Controls:

$$
\hat{w}^{sc} = \underset{w}{\mathrm{argmin}} \ ||\pmb{\bar{y}}_{pre, tr} - \pmb{Y}_{pre, co} \pmb{w}_{co}||^2_2 
$$ $$
\text{s.t } \ \sum w_i = 1 \text{ and } \ w_i > 0 \ \forall \ i
$$

where the outcome $\pmb{Y}_{pre, co}$ is a $T_{pre}$ by $N_{co}$ matrix, where the columns are the units and the rows are the time periods. $\pmb{w}_{co}$ is a $N_{co}$ by 1 column vector, with one entry for each unit. Finally, $\pmb{\bar{y}}_{pre, tr}$ is a $T_{pre}$ by 1 column vector, where each entry is the time average of the treated units in the pre-treatment period.

The **values of synthetic control group** before and after treatment is:

$$
\pmb{y}_{sc} = \pmb{Y}_{co}\hat{\pmb{w}}^{sc}
$$

the average treatment effects ($ATT$) are as follows:

$$
\hat{\tau} =  \bar{y}_{post, tr} - \bar{y}_{post, sc}
$$

Let's say we have a matrix of treatment assignments $D$, where the **columns of the matrix are units** and **rows of the matrix are time periods**.

$$
D = \begin{bmatrix}
    0 & 0 & 0 & \dots & 0 & 0 \\
    0 & 0 & 0 & \dots & 0 & 0 \\
    \vdots \\
    0 & 0 & 0 & \dots & 1 & 1 \\
    0 & 0 & 0 & \dots & 1 & 1 \\
\end{bmatrix}
$$

We merge the equations of DiD and SC and obtain the estimator of SDiD:

$$
\hat{\tau}^{sdid} = \underset{\mu, \alpha, \beta, \tau}{argmin}  \bigg\{ \sum_{i=1}^N \sum_{t=1}^T \big(Y_{it} - (\mu + \alpha_i + \beta_t + \tau D_{it}\big)^2 \hat{w}^{sdid}_i \hat{\lambda}^{sdid}_t \bigg\}
$$

## Results

My preliminary anticipation is that Dobbs' decision make one third of women that would chose to terminate pregnancy change their mind to give birth to a baby.

Non-Parallel Trend

![](images/non-parallel-04.svg){fig-align="center"}

Estimation:

![](images/sdid.svg){fig-align="center"}