### 2025-01-28 ProjectBlog/paper2/DV Adjust/DV Adjust.qmd
p2 dv adjustment < 21:00

---
title: "DV Adjustment"
date: 01/29/2024
date-modified: last-modified
date-format: full
categories: [Paper2]
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
---

Distribution of going to be times.

![](images/Go2Bed-02.png){fig-align="center"}

Distribution of waking up times.

![](images/wake.png){fig-align="center"}