### 2025-02-04 ProjectBlog/paper2/Paper 2 Slides/P2 Slides.qmd
paper 2 slides

---
title: "Social Network Activation: \\The Application in Smartphone Use"
date: 02/04/2025
date-modified: last-modified
date-format: full
categories: [Paper2]
image: "inout.png"
format:
  revealjs: 
    theme: default
smaller: true
bibliography: references.bib
---

### Background

-   Humans are inherently **social beings**. Social engagement improves feeling of belongingness, social fulfillment, etc., which enhance mental health. [@luo2020].

-   In contrast, social isolation can results in anxiety, depression [@robb2020], and suicide [@garnettf.2023].

### **Research Gap**

-   To address the consequences of lack of social engagement, prior studies primarily focus on social interaction volume. However, little attention is paid the the structure of one's social network.

-   Specifically, it is not clear whether activating different parts (e.g., core vs.periphery) of social network impact individuals differently.

### Context:

Social isolation ➟ Smartphone Addiction

##

###

![](inout.png){fig-align="center"}

## Hypothesis 1

-   **H1:** Activating social networks reduces smartphone screen time; the reduction in screen time increases with the number of contacts activated. <br /> <br /> <br />

## Whom to activate?

Identification of Social Influence

-   Social person VS. Non-social person
-   Central VS. Peripheral

## Connectedness

![](images/centrality-01.png){fig-align="center"}

-   **H2:** Activating social networks with a higher **Degree Centrality** score **reduces** the time spent on a smartphone.

## Consider group cohesiveness

![](images/core_number.png){fig-align="center"}

-   **H3:** Activating social networks with a higher **K-Core** score **reduces** the time spent on a smartphone.

## Specification 1

\

$\begin{equation} \begin{split} \text{Screen Time}_{it} = \beta_0 + \beta_1 \cdot \text{Social Network Features}_{it} + \alpha \cdot X_{it} + \theta_i + \lambda_t + \varepsilon_{it} \end{split} \label{e1}\end{equation}$\

-   $\text{Screen Time}_{it}$: Smartphone use time before users fall asleep
-   $\text{Social Network Features}_{it}$: Degree Centrality and K-Core
-   $\text{X}_{it}$: Control variables: Steps, sleep duration, naps, etc.
-   $\theta_i$: Individual fixed effect (students)
-   $\lambda_t$: Time fixed effect (week)

## Result 1 & 2 (Centrality & K-Core):

![](images/2table5.png){fig-align="center"}

## Interpretation

::: incremental
-   Degree Centrality **does not** affect non-communication smartphone use but **improve**s overall smartphone use by increasing the time spent on communication. **H1 is not supported.**
-   K-Core **reduces** overall smartphone use( Communication, others and overall).  **H2 is supported.**
:::

## Mediation of Physical Activity

![](images/paste-1.png){fig-align="center"}

## Sobel-Goodman Mediation Tests

![](/posts/Mediation%20Effect%20of%20Steps/images/paste-6.png){fig-align="center"}

## Mediation of Physical Activity

![](/posts/Mediation%20Effect%20of%20Steps/images/mediation%20-%20before.png){fig-align="center"}

## Interpretation

::: incremental
-   ::: incremental
    Core number reduces screen time by increasing individuals' physical active level

    15.8% of reductive effects of core number come from steps
    :::
:::

## Result 4: Addiction Esimation (Logistic)

\
Addiction is defined as:

-   **Overall** addiction: overall screen time \> 300 min

-   **Non-communication-drive** addiction: Non-communication screen time \> 240 min

-   **Communication-drive** addiction: Communication screen time \> 40 min

## Specification 2

$\begin{equation} \begin{split} \ln \left(\frac{\operatorname{Pr}\left(\text { Addiction }_{it}=1 \mid \boldsymbol{X}_{i t}\right)}{1-\operatorname{Pr}\left(\text { Addiction }_{it}=1 \mid \boldsymbol{X}_{i t}\right)}\right) = \beta_{0}+ \beta_{1} \cdot \text{Social Network Features}_{it}+ \\ \alpha \cdot X_{it}+\theta_{i} +\lambda_{t}+\varepsilon_{it} \end{split} \label{e2} \end{equation}$\

-   $\text{Screen Time}_{it}$: Smartphone use time before users fall asleep
-   $\text{Social Network Features}_{it}$: Degree Centrality and K-Core
-   $\text{X}_{it}$: Control variables: Steps, sleep duration, naps, etc.
-   $\theta_i$: Individual fixed effect (students)
-   $\lambda_t$: Time fixed effect (day)

## Smartphone Addiction Esimation (Logistic)

``` {.python code-line-numbers="12-13"}
global id egoid
global t date2
global ylist addic_noncom90
global mainIV  active_nodes out_contacts  num_edges  steps minsasleep trust close position duration
global Social_Control degree_centrality core_number
global Other_Control i.insession i.weekday community naps nap_minsasleep
global fe i.studyweek
*Set data as panel data
sort $id $t
xtset $id $t

xtreg $ylist $fe  $Other_Control $Social_Control  $mainIV , fe cluster($id) 
xtlogit $ylist $fe  $Other_Control $Social_Control  $mainIV , fe cluster($id) or
outreg2 using SoUnAcSl.tex, append dec(3) addtext(Communication, Full, Unit F.E., YES, Week F.E, YES) keep($mainIV $Social_Control)  title("90 Addiction Estimation")
```

## Result 4: Addiction Esimation (Logistic)

![](images/paste-2.png){fig-align="center"}

##

## Interpretation

::: incremental
-   Network features **outperform** traditional constructs (Trust, Social Distance (Closeness)).
:::

##