### 2025-02-04 ProjectBlog/paper2/Network Paper/p2literature.qmd
is studies on network

---
title: "IS Studies on Social Network"
date: 02/04/2025
date-modified: last-modified
draft: true
date-format: full
categories: [Others]
#image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

# Network Dynamics: How Can We Find Patients Like Us

**Lu (Lucy) Yan, Jianping Peng, Yong Tan**

<iframe src="pdfs/Yan et al. - 2015 - Network Dynamics How Can We Find Patients Like Us.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>