### 2025-02-04 ProjectBlog/paper2/Network Paper/p2literature.qmd
is network

---
title: "IS Studies on Social Network"
date: 02/04/2025
date-modified: last-modified
draft: true
date-format: full
categories: [Others]
#image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

# Network Dynamics: How Can We Find Patients Like Us

**Lu (Lucy) Yan, Jianping Peng, Yong Tan**

<iframe src="pdfs/Yan et al. - 2015 - Network Dynamics How Can We Find Patients Like Us.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

# Feeling Blue? Go Online: An Empirical Study of Social Support Among Patients

**Lu (Lucy) Yan, Yong Tan**

<iframe src="pdfs/Yan and Tan - 2014 - Feeling Blue Go Online An Empirical Study of Social Support Among Patients.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

# Social Networks and the Diffusion of UserGenerated Content: Evidence from YouTube

**Anjana Susarla, Jeong-Ha Oh, Yong Tan**

<iframe src="pdfs/Susarla et al. - 2012 - Social Networks and the Diffusion of User-Generated Content Evidence from YouTube.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

# Network Effects: The Influence of Structural Capital on Open Source Project Success

**Param Vir Singh, Yong Tan, Vijay Mookerjee**

<iframe src="pdfs/Singh et al. - 2011 - Network Effects The Influence of Structural Capital on Open Source Project Success.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

# Location, Location, Location: How Network Embeddedness Affects Project Success in Open Source Systems

**Rajdeep Grewal, Gary L. Lilien, Girish Mallapragada,**
<iframe src="pdfs/Grewal et al. - 2006 - Location, Location, Location How Network Embeddedness Affects Project Success in Open Source System.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>