### 2025-02-11 ProjectBlog/paper2/Paper 2 Slides/P2 Slides.qmd
Extraversion as control

---
title: "Social Network Activation: \\The Application in Smartphone Use"
date: 02/04/2025
date-modified: last-modified
date-format: full
categories: [Paper2]
image: "inout.png"
format:
  revealjs: 
    theme: default
smaller: true
bibliography: references.bib
---

### Motivation

-   Humans are inherently **social beings**. Social engagement improves feeling of belongingness, social fulfillment, etc., which enhance mental health. [@luo2020].

-   In contrast, social isolation can results in anxiety, depression [@robb2020], and suicide [@garnettf.2023].

### **Gap**

-   To address the consequences of lack of social engagement, prior studies primarily focus on social interaction volume.

-   However, little attention is paid the the structure of one's social network. Specifically, it is not clear whether activating different parts (e.g., core vs.periphery) of social network impact individuals differently.

### Study/Methodology

-   We examine whether activating the core parts of social networks, as measured by K-Core, affects smartphone usage activities differently than activating peripheral parts.

##

### Study/Methodology

-   We leverage data sets collected from a series of longitudinal studies that tracked participants’ smartphone use, social networks, and physical activities. We measure social network activation in real-time and track its evolution for each participant, combining self-reported social network structure information and real-time mobile communications.
-   Related studies:

<iframe src="pdfs/Yan et al. - 2015 - Network Dynamics How Can We Find Patients Like Us.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

###

## Hypothesis 1

-   **H1:** Activating social networks reduces smartphone screen time; the reduction in screen time increases with the number of contacts activated. <br /> <br /> <br />

## Specification 1

\

$\begin{equation} \begin{split} \text{Screen Time}_{it} = \beta_0 + \beta_1 \cdot \text{Social Network Features}_{it} + \alpha \cdot X_{it} + \theta_i + \lambda_t + \varepsilon_{it} \end{split} \label{e1}\end{equation}$\

-   $\text{Screen Time}_{it}$: Smartphone use time before users fall asleep
-   $\text{Social Network Features}_{it}$: Degree Centrality and K-Core
-   $\text{X}_{it}$: Control variables: Steps, sleep duration, naps, etc.
-   $\theta_i$: Individual fixed effect (students)
-   $\lambda_t$: Time fixed effect (day)

## Alternative Explanation

![](images/clipboard-3408104519.png){fig-align="center"}

-   Extraverted individuals may incline to activate embeded contacts for group activities.

## Subsample Analysis (Network Size)

![](images/clipboard-2421990295.png){fig-align="center"}

-   K-Core is more effective in reducing screen time for one with small social network**.**

## Subsample Analysis (Steps)

![](images/clipboard-997836528.png){fig-align="center"}

-   K-Core is more effective in reducing screen time for one with high level steps**.**

## Subsample Analysis (Screen Time)

![](images/clipboard-894895312.png){fig-align="center"}

-   K-Core is more effective in reducing screen time for heavy screen users**.**