### 2025-03-04 ProjectBlog/paper2/Paper 2 Slides 2/p2literature.qmd
paper 2 update

---
title: "Slides for Social Network Paper"
date: 02/25/2025
date-modified: last-modified
draft: false
date-format: full
categories: [Paper2]
#image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

# Slides for Social Network Paper

<iframe src="pdfs/Social_Network_Paper_Slides.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

# Outline for Social Network Paper

<iframe src="pdfs/2_27_Paper_2_Outline.pdf" style="width:100%; height:1500px;" frameborder="0">

</iframe>

Conference: https://2025.cswimworkshop.org/