### 2025-04-22 ProjectBlog/paper2/DV Adjust/DV Adjust.qmd

---
title: "DV Adjustment"
date: 01/29/2024
date-modified: last-modified
date-format: full
categories: [Paper2]
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
---

## Distribution of Going to Bed Times

![](images/Go2Bed-02.png){fig-align="center"}

It is not likely that social network activation affects screen time **right before sleep**.

## Distribution of Waking Up Times

![](images/wake.png){fig-align="center"}

## Results with 24 hours Screen Time Measurement

![](images/clipboard-1540611184.png){fig-align="center"}

## Screen Time \< 21:00

![](images/clipboard-2412859547.png){fig-align="center"}

Results with Lagged Core Number

![](images/clipboard-310908795.png){fig-align="center"}

## Potential Mediation Effects

![](images/clipboard-4063857286.png){fig-align="center"}