### 2025-04-22 ProjectBlog/paper2/Paper 2/paper2.qmd

---
title: "Social Network Activation: \\The Application in Smartphone Use"
date: 10/09/2023
date-modified: last-modified
date-format: full
categories: [Paper2]
image: "inout.png"
format:
  revealjs: 
    theme: default
smaller: true
bibliography: references.bib
---

### Background

Humans are inherently **social beings**.

-   Social engagement ➟ belongingness, social fulfillment, etc [@luo2020].

-   Social isolation ➟ anxiety, depression [@robb2020], and suicide [@garnettf.2023].

### **Research Gap**

-   Social interaction volume VS. **Structure/Topology** ➟ Behavior/Health?

-   Social network **activation/evolution** ➟ Behavior/Health?

### Context:

Social isolation ➟ Smartphone Addiction

##

###

![](inout.png){fig-align="center"}

## Hypothesis 1

-   **H1:** Activating social networks reduces smartphone screen time; the reduction in screen time increases with the number of contacts activated. <br /> <br /> <br />

## Whom to activate?

Identification of Social Influence

-   Social person VS. Non-social person
-   Central VS. Peripheral

## Connectedness

![](images/centrality-01.png){fig-align="center"}

-   **H2:** Activating social networks with a higher **Degree Centrality** score **reduces** the time spent on a smartphone.

## Consider group cohesiveness

![](images/core_number.png){fig-align="center"}

-   **H3:** Activating social networks with a higher **K-Core** score **reduces** the time spent on a smartphone.

## Methodology

<br> ![](images/method.png){fig-align="center"}

## How SNA features are calculated

![Survey t-1](images/Demo-t-1-2.png){fig-align="center"}

## How SNA features are calculated

![Survey t-1](images/Demo-t-1.png){fig-alt="The normalized degree centrality for a node in an undirected network is calculated by dividing the degree of the node (the number of its connections) by the maximum possible degree in the network. The maximum possible degree is N−1, where N is the total number of nodes in the network." fig-align="center"}

| Node        | Connections | Max Degree N−1 | Degree Centrality |
|-------------|-------------|----------------|-------------------|
| A           | 2           | 3              | 2/3               |
| B           | 2           | 3              | 2/3               |
| C           | 3           | 3              | 1                 |
| Participant | 3           | 3              | 1                 |

## How SNA features are calculated

![Survey t](images/Demo-t.png){fig-align="center"}

## How SNA features are extracted

![Survey t+1](images/Demo-t+1.png){fig-align="center"}

## Specification 1

\

$\begin{equation} \begin{split} \text{Screen Time}_{it} = \beta_0 + \beta_1 \cdot \text{Social Network Features}_{it} + \alpha \cdot X_{it} + \theta_i + \lambda_t + \varepsilon_{it} \end{split} \label{e1}\end{equation}$\

-   $\text{Screen Time}_{it}$: Smartphone use time before users fall asleep
-   $\text{Social Network Features}_{it}$: Degree Centrality and K-Core
-   $\text{X}_{it}$: Control variables: Steps, sleep duration, naps, etc.
-   $\theta_i$: Individual fixed effect (students)
-   $\lambda_t$: Time fixed effect (week)

## Result 1 & 2 (Centrality & K-Core):

![](images/2table5.png){fig-align="center"}

## Interpretation

::: incremental
-   Degree Centrality **does not** affect non-communication smartphone use but **improve**s overall smartphone use by increasing the time spent on communication. **H1 is not supported.**
-   K-Core **reduces** overall smartphone use( Communication, others and overall).  **H2 is supported.**
:::

## Mediation of Physical Activity

![](images/paste-1.png){fig-align="center"}

## Sobel-Goodman Mediation Tests

![](/posts/Mediation%20Effect%20of%20Steps/images/paste-6.png){fig-align="center"}

## Mediation of Physical Activity

![](/posts/Mediation%20Effect%20of%20Steps/images/mediation%20-%20before.png){fig-align="center"}

## Interpretation

::: incremental
-   ::: incremental
    Core number reduces screen time by increasing individuals' physical active level

    15.8% of reductive effects of core number come from steps
    :::
:::

## Result 4: Addiction Esimation (Logistic)

\
Addiction is defined as:

-   **Overall** addiction: overall screen time \> 300 min

-   **Non-communication-drive** addiction: Non-communication screen time \> 240 min

-   **Communication-drive** addiction: Communication screen time \> 40 min

## Specification 2

$\begin{equation} \begin{split} \ln \left(\frac{\operatorname{Pr}\left(\text { Addiction }_{it}=1 \mid \boldsymbol{X}_{i t}\right)}{1-\operatorname{Pr}\left(\text { Addiction }_{it}=1 \mid \boldsymbol{X}_{i t}\right)}\right) = \beta_{0}+ \beta_{1} \cdot \text{Social Network Features}_{it}+ \\ \alpha \cdot X_{it}+\theta_{i} +\lambda_{t}+\varepsilon_{it} \end{split} \label{e2} \end{equation}$\

-   $\text{Screen Time}_{it}$: Smartphone use time before users fall asleep
-   $\text{Social Network Features}_{it}$: Degree Centrality and K-Core
-   $\text{X}_{it}$: Control variables: Steps, sleep duration, naps, etc.
-   $\theta_i$: Individual fixed effect (students)
-   $\lambda_t$: Time fixed effect (day)

## Smartphone Addiction Esimation (Logistic)

``` {.python code-line-numbers="12-13"}
global id egoid
global t date2
global ylist addic_noncom90
global mainIV  active_nodes out_contacts  num_edges  steps minsasleep trust close position duration
global Social_Control degree_centrality core_number
global Other_Control i.insession i.weekday community naps nap_minsasleep
global fe i.studyweek
*Set data as panel data
sort $id $t
xtset $id $t

xtreg $ylist $fe  $Other_Control $Social_Control  $mainIV , fe cluster($id) 
xtlogit $ylist $fe  $Other_Control $Social_Control  $mainIV , fe cluster($id) or
outreg2 using SoUnAcSl.tex, append dec(3) addtext(Communication, Full, Unit F.E., YES, Week F.E, YES) keep($mainIV $Social_Control)  title("90 Addiction Estimation")
```

## Result 4: Addiction Esimation (Logistic)

![](images/paste-2.png){fig-align="center"}

##

## Interpretation

::: incremental
-   Network features **outperform** traditional constructs (Trust, Social Distance (Closeness)).
:::

##