### 2025-08-17 index.qmd
job website

---
title: "Shaokang Yang"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

Department of Business Information Technology

Pamplin College of Business, Virginia Tech

Blacksburg, VA 24061

Website: [shaokang-yang.net](shaokang-yang.net)

-   **Virginia Polytechnic Institute and State University**

    Ph.D. in Business Information Technology (2026, Expected)

    M.A. in Data Analysis and Applied Statistics

-   **University of Milan**

    M.Sc. in Data Science and Economics

-   **Jimei University**

    Bachelor of Engineering; Bachelor of Economics