### 2025-08-17 research.qmd

---
title: "Research"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

## Research Interests

-   Topics: Healthcare policy analysis, healthcare IT
-   Methodologies: Causal inference with econometrics, machine learning, statistical learning

## Working Papers

-   **Smartphone Use, Social Support, and Sleep Health** Shaokang Yang, Idris Adjerid. \[[SSRN](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=5368531)\] –*Under Review at Information Systems Research*

-   **Friends over Feeds: How Social Smartphone Use Reduces Screen Time** Shaokang Yang, Idris Adjerid. – *In preparation for submission*\

<!-- -->

-   ***The Abortion War: When Restrictions Meet Technology** Shaokang Yang, Idris Adjerid.* – *In preparation for submission*

## Presentations

-   **Friends over Feeds: How Social Smartphone Use Reduces Screen Time.**\
    Shaokang Yang, Idris Adjerid
    -   Conference on Health IT and Analytics (CHITA) — Apr 2025, Austin, TX
    -   INFORMS Annual Meeting — Oct 2024, Seattle, WA
-   **Smartphone Use, Social Support, and Sleep Health.**\
    Shaokang Yang, Idris Adjerid
    -   Conference on Health IT and Analytics (CHITA) — May 2024, Washington, DC
    -   33rd Annual POMS Conference — May 2023, Orlando, FL
    -   Conference on Health IT and Analytics (CHITA) — May 2023, Washington, DC