### 2025-08-17 teaching.qmd

---
title: "Teaching"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

## Instructor

-   **BIT 3424 — Intro to Business Analytics Modelling** — Summer 2024

*Undergraduate advanced course; programming with VBA, Python, and R*

## Tutorial Instructor

-   **BIT 4624 — Cybersecurity Analytics** — Fall 2022; Spring 2023; Fall 2024

*Taught tutorials of Cybersecurity Analytics (machine learning) with R*