### 2025-09-01 research.qmd
research update

---
title: "Research"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

## Research Interests

-   Topics: Policy evaluation on health IT, Digital technologies, Economics of AI
-   Methodologies: Causal inference with econometrics, machine learning, statistical learning

## Working Papers

### Digital Technologies and Well-being

1. **Smartphone Use, Social Support, and Sleep Health.** Shaokang Yang, Idris Adjerid. [SSRN](http://dx.doi.org/10.2139/ssrn.5368531) -- *Under Review at Information Systems Research*

2. **Friends over Feeds: How Social Smartphone Use Reduces Screen Time.** Shaokang Yang, Idris Adjerid. -- *In preparation for submission (target: Organization Science)*

3. **Pet Ownership and Mental Health During the Pandemic: Evidence from an E-Commerce Platform** Li Lin, Shaokang Yang -- *Data collection phase*

### Policy Evaluation on Health IT

1. **The Abortion War: When Restrictions Meet Technology.** Shaokang Yang, Idris Adjerid. -- *Data collected, working on empirical analysis*

2. **Pro-Choice Vs Pro-Life: The Impact of Telehealth Abortion on Infant Health.** -- *Data collected, working on empirical analysis*

## Presentations

-   **Friends over Feeds: How Social Smartphone Use Reduces Screen Time.**\
    Shaokang Yang, Idris Adjerid
    -   Conference on Health IT and Analytics (CHITA) — Apr 2025, Austin, TX
    -   INFORMS Annual Meeting — Oct 2024, Seattle, WA
-   **Smartphone Use, Social Support, and Sleep Health.**\
    Shaokang Yang, Idris Adjerid
    -   Conference on Health IT and Analytics (CHITA) — May 2024, Washington, DC
    -   33rd Annual POMS Conference — May 2023, Orlando, FL
    -   Conference on Health IT and Analytics (CHITA) — May 2023, Washington, DC