### 2025-09-03 index.qmd
website

---
title: "Shaokang Yang"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

<table style="width:100%">
  <tr>
    <td>Department of Business Information Technology</td>
    <td align="right">Tel: 540-824-8612</td>
  </tr>
  <tr>
    <td>Pamplin College of Business, Virginia Tech</td>
    <td align="right">Email: shaokang@vt.edu</td>
  </tr>
  <tr>
    <td>Blacksburg, VA 24061</td>
    <td align="right"><a href="https://wholehealth.isce.vt.edu/members-directory/yang-shaokang-bio.html">Whole Health Consortium</a></td>
  </tr>
</table>

------------------------------------------------------------------------

**Virginia Polytechnic Institute and State University**

-   Ph.D. in Business Information Technology (2026, Expected)

-   M.A. in Data Analysis and Applied Statistics

**University of Milan**

-   M.Sc. in Data Science and Economics

**Jimei University**

-   Bachelor of Engineering; Bachelor of Economics