### 2025-09-17 ProjectBlog/paper1/Pre3HoursSleep/chita.qmd

---
title: "Screen Time Before Sleep"
date: 04/22/2025
date-modified: last-modified
date-format: full
categories: [Paper1]
#image: "search.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

# Measurement Window

![](images/clipboard-727691413.png){fig-align="center"}

## Descriptive Statistics

```{python}
#| echo: false
#| warning: false
#| output: true
import time;import pandas as pd;import numpy as np;import datetime
from datetime import datetime;from datetime import timedelta;import sqlite3;import warnings
import matplotlib.pyplot as plt;import seaborn as sns;warnings.filterwarnings('ignore')

df=pd.read_csv("/Users/shaokangyang/Library/CloudStorage/GoogleDrive-sky.ang510@gmail.com/My Drive/Nethealth/Data/CNSA/combine/7/df_pre_3hr.csv")

df[['major_minstofallasleep', 'sleep_duration','major_minsawake','sleep_debt', 
    'major_minsafterwakeup',  'major_bedtime',]].describe().round(2)
```

## Model-Free Evidence

![](images/clipboard-3316574643.png){fig-align="center"}

## Waking Up Time as DV

![](images/clipboard-3013034043.png){fig-align="center"}

## Confidence Intervals Plot

![](images/clipboard-3836821377.png){fig-align="center"}

![](images/clipboard-2494050607.png){fig-align="center"}

# CHTIA

The slides for CHITA 2024 are below:

<iframe src="pdfs/Social_Network_Paper_Slides___Yang_Adjerid_Liu.pdf" style="width:150%; height:1500px;" frameborder="0">

</iframe>

You can download it [here](pdfs/CHITA2024.pdf).