### 2025-09-17 index.qmd

---
title: "Shaokang Yang"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

+-----------------------------------------------+-----------------------------------------------------------------------------------------------------+
| Department of Business Information Technology | Tel: 540-824-8612                                                                                   |
+-----------------------------------------------+-----------------------------------------------------------------------------------------------------+
| Pamplin College of Business, Virginia Tech    | Email: shaokang\@vt.edu                                                                             |
+-----------------------------------------------+-----------------------------------------------------------------------------------------------------+
| Blacksburg, VA 24061                          | [Whole Health Consortium](https://wholehealth.isce.vt.edu/members-directory/yang-shaokang-bio.html) |
+-----------------------------------------------+-----------------------------------------------------------------------------------------------------+

------------------------------------------------------------------------

**Virginia Polytechnic Institute and State University**

-   Ph.D. in Business Information Technology (2026, Expected)

-   M.A. in Data Analysis and Applied Statistics

**University of Milan**

-   M.Sc. in Data Science and Economics

**Jimei University**

-   Bachelor of Engineering; Bachelor of Economics