### 2025-09-17 posts/Daylight Saving Time/daylight.qmd

---
title: "Daylight saving time"
date: 04/23/2024
date-modified: last-modified
date-format: full
categories: [Paper1]
image: "time.png"
tbl-colwidths: [5,50,75]
format: 
  html:
    page-layout: full
---

In this blog, we check our results could be impacted by daylight saving time.Daylight saving time refers to the practice of setting the clocks forward by one hour from standard time during the warmer months of the year, in order to make better use of natural daylight. The practice is used in many European countries, as well as in the United States. The idea is that by moving the clocks forward by one hour, people can enjoy more daylight in the evenings, and this can help to reduce energy consumption and promote outdoor activities. Since the data sets we use are longitudinal studies, we need to consider the impact of daylight saving time on our results. The dates of daylight saving time change in this study are as follows: **"2015-03-08", "2015-11-01", "2016-03-13", "2016-11-06", "2017-03-12", "2017-11-05", "2018-03-11", "2018-11-04"**.

### Anomalies Caused by Daylight saving time Change

The following table shows the smartphone use activities for egoid "79897" on a daylight saving time change date **"2016-11-06".** The measurement window for the day range from **2016-11-05 22:22:00** to **2016-11-06 01:22:00,** the observations between the two yellow rows are the those captured by our measurement.

![](images/clipboard-3850785954.png){fig-align="center"}

We observe that the clocks are set backward by one hour from daylight saving time to standard time (orange cells). If there were no daylight saving time change, the ending point of the measurement window will be **2016-11-06 02:22:00. In another word, the three-hour** measurement window is extended to **four** hours on this day. However, since the two observations at 1:57 and 1:59 are out of the measurement window, they are **not** considered in the aggregation process. Therefore, the daylight saving time change caused some anomalies of screen time and unlock.

### Impact of Time Adjustment on Sleep

Due to the clocks change in the Fall and Spring (Spring Forward，Fall Back). Individuals have one hour less sleep on spring change day, and one hour more sleep on fall. Therefore, we may need to drop the observation on those days.