### 2025-09-17 posts/Survey Overlap/index.qmd

---
title: "Network Overlapping"
date: 01/29/2024
date-modified: last-modified
date-format: full
categories: [Paper2]
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
---

Since participants report their network contacts over social network survey, some contacts may be repteatedly reprted in multiple surveys. To understand whether students' social network survey are cumulative or aggregated over time, in this blog, we exmaine the overlapping rate of reported contacts.

The following figure shows how many waves of surveys participants have ever joined.

![](images/paste-1.png)

For example, 78 participants completed 6 surveys.

```{python}
#| echo: false
#| warning: false
#| output: false
import time;import pandas as pd;import numpy as np;import datetime
from datetime import datetime;from datetime import timedelta;import sqlite3
import warnings;warnings.filterwarnings('ignore')
vt=pd.read_csv("/Users/shaokangyang/Library/CloudStorage/OneDrive-Personal/Nethealth/Data/overlap_rate.csv")
```

The following table shows the social network summary of participant "10237".

```{python}
#| echo: false
#| warning: false
#| output: true
vt10237 = vt[vt['egoid'] == 10237]
vt10237 = vt10237.drop(columns=['vertex1', 'vertex2'])
print(vt10237.to_string(index=False))
```

-   **egoid:** individual identifier;

-   **wave:** survey identifier;

-   **surveys:** the number of surveys the participant completed;

-   **contacts:** the number of contacts reported in the current survey;

-   **total contacts:** the number of contacts reported in total survey;

-   **participation_rate:** contacts/total contacts;

-   **contact_ids:** the identifiers of contacts reported in the current survey;

-   **overlap:** the overlapping rate of contacts of previous and current survey, which is calculated as: the number of contacts reported both by the previous and current surveys / the number of contacts in the current survey.

In the survey of wave 2, contacts "**395167(Sibling), 791455(Sibling)**" has been reported in the survey of wave 1, the overlap rate is 2/6=0.33.

In the survey of wave 3, contacts **"760234"(Romantic Partner)** has been reported in the survey of wave 2, the overlap rate is 1/13=0.076.

In the survey of wave 4, contacts "**982757(Sibling), 390853(Friend), 899046(Stepsibling/Friend), 349320(Friend), 760234(Romantic Partner), 857038(Romantic Partner/OtherFamily), 609614(Parent)**" has been reported in the survey of wave 3, the overlap rate is 7/10=0.70.

In the survey of wave 5, contacts " **760234(Romantic Partner), 390853(Friend), 899046(Stepsibling/Friend),, 277215(Friend)**" has been reported in the survey of wave 4, the overlap rate is 4/16=0.25.

In the survey of wave 6, contacts " **669474(Friend), 390853(Friend), 760234(Romantic Partner), 438507(Friend), 549070(Parent), 277215(Friend), 791455(Sibling), 395167(Sibling), 193183(Friend)**" has been reported in the survey of wave 5, the overlap rate is 9/15=0.60.

In the survey of wave 7, contacts " **760234(Romantic Partner), 486534(Friend), 193183(Friend)**" has been reported in the survey of wave 6, the overlap rate is 3/6=0.50.

The following table shows the descriptive summary for all participants (Person $⨯$ Wave of Survey)

```{python}
#| echo: false
#| warning: false
#| output: true
vt.describe()
```