### 2025-10-09 teaching.qmd

---
title: "Teaching"
format:
  html:
    css: styles2.css
class: custom-font-pos
---

## Teaching Interests

-   Data analytics, machine learning, and AI-powered programming

-   Cybersecurity analytics, healthcare analytics and policy evaluation

-   Econometrics for business research

## Instructor

-   **BIT 3424 — Intro to Business Analytics Modelling** — Summer 2024

*Undergraduate advanced course; programming with VBA, Python, and R*

## Tutorial Instructor

-   **BIT 4624 — Cybersecurity Analytics**

Cybersecurity tutorials in data analytics and machine learning with R — Fall 2025

Cybersecurity tutorials in data analytics and machine learning with R — Fall 2024

Cybersecurity tutorials in data analytics and machine learning with R — Spring 2023

Cybersecurity tutorials in data analytics and machine learning with R — Fall 2022